Cherenkov radiation and scattering of external dispersive waves by two-color solitons (Supplemental Material)

### 2022-07-07 text/Supplementary.tex

\documentclass[a4paper,11pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\newcommand{\abs}[1]{\left| #1 \right|}

\title{Supplementary Material}
\date{\today}

@@

@@

@@

When analyzing the nonlinear scattering near an oscillatory mode we have used an expression for the oscillation frequency. In this section we will derive this expression.

Let us return to equation (3) for coupled solitons. We can re-normalize the equations by performing the following transformation
\begin{equation*}
  U_{n} \to \gamma_{n}^{1/2} e^{i \beta_{n} z} \cdot U_{n},
\end{equation*}
which will make the equations symmetric
\begin{equation*}
  i \partial_{z} U_{n}
    - \frac{1}{2} \beta''_{n} \partial_{t}^{2} U_{n}
    + \gamma_{n}^{2} \abs{U_{n}}^{2} U_{n}
    + 2 \gamma_{n} \gamma_{m} \abs{U_{m}}^{2} U_{n} = 0.
\end{equation*}
This in turn allows us to recognize the modified couple of equations as Euler-Lagrange equations for Lagrangian
\begin{equation*}
  \int \limits_{-\infty}^{+\infty}
    \mathcal{L}(U_{1}, \partial_{z} U_{1}, \partial_{t} U_{1}, \ldots)
    \, dt,
\end{equation*}
where Lagrangian density $\mathcal{L}$ is defined as a sum of three components $\mathcal{L} = \mathcal{L}_{1} + \mathcal{L}_{2} + \mathcal{L}_{\text{int}}$, with $L_{n}$ being a single-soliton Lagrangian density
\begin{equation}
  \label{eq:SingleSolitonLagrangianDensity}
  \mathcal{L}_{n} =
  \frac{i}{2} \left(
    \partial_{z} U_{n} \cdot U_{n}^{*} -
    \partial_{z} U_{n}^{*} \cdot U_{n}
  \right)
  + \frac{1}{2} \beta''_{n} \, \partial_{t} U_{n} \partial_{t} U_{n}^{*}
  + \frac{1}{2} \gamma_{n}^{2} \, \abs{U_{n}}^{4},
\end{equation}
and $\mathcal{L}_{\text{int}}$ being the interaction term
\begin{equation}
  \label{eq:InteractionLagrangianDensity}
  \mathcal{L}_{\text{int}} =
    2 \gamma_{1} \gamma_{2}
    \abs{U_{1}}^{2} \abs{U_{2}}^{2}.
\end{equation}
Let us assume that the soliton components $U_{n}$ can be described by the following generic \textit{ansatz}
\begin{equation}
  \label{eq:SolitonAnsatz}
  U_{n}(z, t) = A_{n}(z) S\left(
    \frac{t - t_{n}(z)}{\sigma_{n}(z)}
  \right) \exp\left(
    - i \Omega_{n}(z) t + i \phi_{n}(z)
  \right).
\end{equation}
In here $A_{n}$ is the amplitude of the pulse, $t_{n}$ is the central position, $\sigma_{n}$ is the pulse width, $\Omega_{n}$ is the frequency detuning, $\phi_{n}$ is the phase, and $S(x)$ is function that defines the envelope shape. At the moment we will not specify the concrete form of $S(x)$, but will assume that it is an even function. Substituting \eqref{eq:SolitonAnsatz} into \eqref{eq:SingleSolitonLagrangianDensity} and \eqref{eq:InteractionLagrangianDensity} and integrating over $t$ we arrive at the expressions for the averaged Lagrangians
\begin{gather}
  L_{n} =
    I_{1} \, t_{n} \sigma_{n} A_{n}^{2} \frac{d \Omega_{n}}{d z} +
    I_{1} \, \sigma_{n} A_{n}^{2} \frac{d \phi_{n}}{d z} +
    I_{2} \, \frac{\beta''_{n}}{2} \frac{A_{n}^{2}}{\sigma_{n}} \nonumber \\
    + I_{1} \, \frac{\beta''_{n}}{2} \sigma_{n} A_{n}^{2} \Omega_{n}^{2} +
    I_{3} \, \frac{\gamma_{n}^{2}}{2} \sigma_{n} A_{n}^{4} \\
  L_{\text{int}} =
    2 \, \gamma_{1} \gamma_{2} \, A_{1}^{2} A_{2}^{2} \,
    I_{\text{int}}(\sigma_{1}, \sigma_{2}, t_{1}, t_{2}),
\end{gather}
where the following integrals have been defined
\begin{align*}
  I_{1} =
    \int \limits_{-\infty}^{+\infty}
    S^{2}(x) \, dx &&
  I_{2} =
    \int \limits_{-\infty}^{+\infty}
    (S'(x))^{2} \, dx \\
  I_{3} =
    \int \limits_{-\infty}^{+\infty}
    S^{4}(x) \, dx &&
  I_{\text{int}} =
    \int \limits_{-\infty}^{+\infty}
    S^{2}\left(
      \frac{t - t_{1}}{\sigma_{1}}
    \right)
    S^{2}\left(
      \frac{t - t_{2}}{\sigma_{2}}
    \right) \, dt
\end{align*}
Due to the time invariance in the problem, $I_{\text{int}}$ depends only on the difference between $t_{1}$ and $t_{2}$
\begin{equation*}
  I_{\text{int}} = I_{\text{int}}(t_{1} - t_{2}, \sigma_{1}, \sigma_{2}),
\end{equation*}
and it is an even function of that difference.

The averaged Lagrangian $L = L_{1} + L_{2} + L_{\text{int}}$ is now a function defined in terms of soliton parameters \{\,$A_{n}$, $\sigma_{n}$, $t_{n}$, $\Omega_{n}$, $\phi_{n}$\,\} and only them. Therefore, the Euler-Lagrange equations for the new Lagrangian have to be defined in terms of variations over the soliton parameters
\begin{equation*}
  \frac{\delta L}{\delta P_{n}} =
    \frac{\partial{L}}{{\partial P_{n}}} -
    \frac{d}{dz} \frac{\partial L}{\partial \dot P_{n}} = 0,
\end{equation*}
where $P_{n}$ stands for either $A_{n}$, $\sigma_{n}$, $t_{n}$, $\Omega_{n}$ or $\phi_{n}$. The latter case --- variation with respect to the phase $\phi_{n}$ --- immediately yields the conservation of mass
\begin{equation}
  \label{eq:ConservationOfMass}
  N_{n} = \sigma_{n}(z) A_{n}^{2}(z) = const.
\end{equation}

@@

Variation with respect to the soliton position $t_{n}$ gives us an equation for the frequency
\begin{equation}
  \label{eq:SolitonFrequency}
  \frac{d \Omega_{n}}{dz} =
    2 \cdot \frac{N_{m} \gamma_{1} \gamma_{2}}
           {I_{1} \sigma_{1}(z) \sigma_{2}(z)} \cdot
    \frac{\partial I_{\text{int}}}{\partial t_{n}}.
\end{equation}
The symmetry in the overlap integral $I_{\text{int}}$ with respect to the soliton positions $t_{1}$ and $t_{2}$ leads to conservation of momentum
\begin{equation}
  \label{eq:ConservationOfMomentum}
  N_{1} \Omega_{1}(z) + N_{2} \Omega_{2}(z) = const.
\end{equation}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\documentclass[aps, pra, signlecolumn, superscriptaddress, final, 11pt]{revtex4}
\pdfoutput=1

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{tikz}

\newcommand{\abs}[1]{\left| #1 \right|}
\renewcommand{\theequation}{S.\arabic{equation}}
\renewcommand{\thefigure}{S.\arabic{figure}}

@@

\title{Cherenkov radiation and scattering of external dispersive waves by two-color solitons (Supplemental Material)}

\author{Ivan Oreshnikov}
\affiliation{Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 T\"ubingen, Germany}

\author{Oliver Melchert}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Stephanie Willms}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}

\author{Surajit Bose}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Photonics, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Ihar Babushkin}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}

\author{Ayhan Demircan}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Uwe Morgner}
\affiliation{Cluster of Exellence PhoenixD, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Leibniz Universit\"at Hannover, Institute of Quantum Optics, Welfengarten 1, 30167 Hannover, Germany}
\affiliation{Hannover Centre for Optical Technologies, Nienburger Strasse 17, 30167 Hannover, Germany}

\author{Alexey Yulin}
\affiliation{Department of Physics, ITMO University, Kronverskiy pr. 49, 19701 St.~Petersburg, Russia}

\date{\today}

@@

\section*{Dispersion profile model}

We model dispersion coefficient $\beta(\omega)$ with the following rational expression
\begin{equation}
  \label{eq:Beta}
  \beta(\omega) = \frac{1}{c} \frac{
    \sum_{n=0}^{3} {C_{n} \omega^{n+1}}
  }{
    \sum_{m=0}^{3} {D_{m} \omega^{m}}
    }
\end{equation}
where $c = 0.299792458~\mu
\text{m/fs}$ is the speed of light, and the coefficient sequences $C$ and $D$ are
defined by
\begin{eqnarray}
C &=& (9.654, -39.739 \,\mathrm{fs},  16.885 \,\mathrm{fs^2}, -2.746 \,\mathrm{fs^3}),\quad \text{and},\\
D &=& (1, -9.496 \,\mathrm{fs}, 4.221 \,\mathrm{fs^2}, -0.703 \,\mathrm{fs^3}).
\end{eqnarray}
%
In here and everywhere in the paper we assume fs as a unit of
time and $\mu$m as a unit of distance. Figure \ref{fig:Model} displays group
velocity $v_{g}(\omega) = 1 / \beta'(\omega)$ and second order dispersion
coefficient $\beta''(\omega)$ as functions of frequency. Frequencies
$\omega_{1}$ and $\omega_{2}$ correspond to the central frequencies of the
soliton's spectral components as chosen in the simulation corresponding to
Fig.\,1 of the paper.

\begin{figure}[ht]
  \centering
  \includegraphics{{Figures/FigA}.pdf}
  \caption{
    (a) group velocity $v_{g}$ and (b) second order dispersion coefficient $\beta''(\omega)$ in the model fiber. Labels $A_{1,2}$ and $N$ mark the regions of anomalous and normal dispersion.
  }
  \label{fig:Model}
\end{figure}

@@

When analyzing the nonlinear scattering near an oscillatory mode we used an expression for the oscillation frequency. In this section we will derive this expression.

Let us return to Eq.~(3) (of the main text) for coupled solitons. We can re-normalize the equations by performing the following transformation
\begin{equation*}
  U_{n} \to \gamma_{n}^{1/2} e^{i \beta_{n} z} \cdot U_{n},
\end{equation*}
which will make the equations symmetric
\begin{equation*}
  i \partial_{z} U_{n}
    - \frac{1}{2} \beta''_{n} \partial_{t}^{2} U_{n}
    + \gamma_{n}^{2} \abs{U_{n}}^{2} U_{n}
    + 2 \gamma_{n} \gamma_{m} \abs{U_{m}}^{2} U_{n} = 0.
\end{equation*}
This in turn allows us to recognize the modified couple of equations as Euler-Lagrange equations for Lagrangian
\begin{equation*}
  \int \limits_{-\infty}^{+\infty}
    \mathcal{L}(U_{1}, \partial_{z} U_{1}, \partial_{t} U_{1}, \ldots)
    \, dt,
\end{equation*}
where Lagrangian density $\mathcal{L}$ is defined as a sum of three components $\mathcal{L} = \mathcal{L}_{1} + \mathcal{L}_{2} + \mathcal{L}_{\text{int}}$, with $\mathcal{L}_{n}$ being a single-soliton Lagrangian density
\begin{equation}
  \label{eq:SingleSolitonLagrangianDensity}
  \mathcal{L}_{n} =
  \frac{i}{2} \left(
    \partial_{z} U_{n} \cdot U_{n}^{*} -
    \partial_{z} U_{n}^{*} \cdot U_{n}
  \right)
  + \frac{1}{2} \beta''_{n} \, \partial_{t} U_{n} \partial_{t} U_{n}^{*}
  + \frac{1}{2} \gamma_{n}^{2} \, \abs{U_{n}}^{4},
\end{equation}
and $\mathcal{L}_{\text{int}}$ being the interaction term
\begin{equation}
  \label{eq:InteractionLagrangianDensity}
  \mathcal{L}_{\text{int}} =
    2 \gamma_{1} \gamma_{2}
    \abs{U_{1}}^{2} \abs{U_{2}}^{2}.
\end{equation}
Let us assume that the soliton components $U_{n}$ can be described by the following generic \textit{ansatz}
\begin{equation}
  \label{eq:SolitonAnsatz}
  U_{n}(z, t) = A_{n}(z) S\left(
    \frac{t - t_{n}(z)}{\sigma_{n}(z)}
  \right) \exp\left(
    - i \Omega_{n}(z) t + i \phi_{n}(z)
  \right).
\end{equation}
In here $A_{n}$ is the amplitude of the pulse, $t_{n}$ is the central position, $\sigma_{n}$ is the pulse width, $\Omega_{n}$ is the frequency detuning, $\phi_{n}$ is the phase, and $S(x)$ is function that defines the envelope shape. At the moment we will not specify the concrete form of $S(x)$, but will assume that it is an even function.

Before we continue let us stress one important thing: this ansatz cannot express all the possible internal oscillations of the soliton. One obvious example, as it was noted in the text, is the case of the pulse-width oscillation. In order to capture this dynamics, we need to add frequency chirp to the ansatz.

Substituting \eqref{eq:SolitonAnsatz} into \eqref{eq:SingleSolitonLagrangianDensity} and \eqref{eq:InteractionLagrangianDensity} and integrating over $t$ we arrive at the expressions for the averaged Lagrangians
\begin{gather}
  L_{n} =
    I_{1} \, t_{n} \sigma_{n} A_{n}^{2} \frac{d \Omega_{n}}{d z} +
    I_{1} \, \sigma_{n} A_{n}^{2} \frac{d \phi_{n}}{d z} +
    I_{2} \, \frac{\beta''_{n}}{2} \frac{A_{n}^{2}}{\sigma_{n}} \nonumber \\
    + I_{1} \, \frac{\beta''_{n}}{2} \sigma_{n} A_{n}^{2} \Omega_{n}^{2} +
    I_{3} \, \frac{\gamma_{n}^{2}}{2} \sigma_{n} A_{n}^{4} \\
  L_{\text{int}} =
    2 \, \gamma_{1} \gamma_{2} \, A_{1}^{2} A_{2}^{2} \,
    I_{\text{int}}(\sigma_{1}, \sigma_{2}, t_{1}, t_{2}),
\end{gather}
where the following integrals have been defined
\begin{align*}
  I_{1} =
    \int \limits_{-\infty}^{+\infty}
    S^{2}(x) \, dx &&
  I_{2} =
    \int \limits_{-\infty}^{+\infty}
    (S'(x))^{2} \, dx \\
  I_{3} =
    \int \limits_{-\infty}^{+\infty}
    S^{4}(x) \, dx &&
  I_{\text{int}} =
    \int \limits_{-\infty}^{+\infty}
    S^{2}\left(
      \frac{t - t_{1}}{\sigma_{1}}
    \right)
    S^{2}\left(
      \frac{t - t_{2}}{\sigma_{2}}
    \right) \, dt
\end{align*}
Due to the time invariance in the problem, $I_{\text{int}}$ depends only on the difference between $t_{1}$ and $t_{2}$
\begin{equation*}
  I_{\text{int}} = I_{\text{int}}(t_{1} - t_{2}, \sigma_{1}, \sigma_{2}),
\end{equation*}
and it is an even function of that difference.

The averaged Lagrangian $L = L_{1} + L_{2} + L_{\text{int}}$ is now a function defined in terms of soliton parameters \{\,$A_{n}$, $\sigma_{n}$, $t_{n}$, $\Omega_{n}$, $\phi_{n}$\,\} and only them. Therefore, the Euler-Lagrange equations for the new Lagrangian have to be defined in terms of variations over the soliton parameters
\begin{equation*}
  \frac{\delta L}{\delta P_{n}} =
    \frac{\partial{L}}{{\partial P_{n}}} -
    \frac{d}{dz} \frac{\partial L}{\partial \dot P_{n}} = 0,
\end{equation*}
where $P_{n}$ stands for either $A_{n}$, $\sigma_{n}$, $t_{n}$, $\Omega_{n}$ or $\phi_{n}$. The latter case --- variation with respect to the phase $\phi_{n}$ --- immediately yields the conservation of mass
\begin{equation}
  \label{eq:ConservationOfMass}
  N_{n} = \sigma_{n}(z) A_{n}^{2}(z) = \mathrm{const.}
\end{equation}

@@

Variation with respect to the soliton position $t_{n}$ gives us an equation for the frequency
\begin{equation}
  \label{eq:SolitonFrequency}
  \frac{d \Omega_{n}}{dz} =
    2 \cdot \frac{N_{m} \gamma_{1} \gamma_{2}}
           {I_{1} \sigma_{1}(z) \sigma_{2}(z)} \cdot
    \frac{\partial I_{\text{int}}}{\partial t_{n}}.
\end{equation}
The symmetry in the overlap integral $I_{\text{int}}$ with respect to the soliton positions $t_{1}$ and $t_{2}$ leads to conservation of momentum
\begin{equation}
  \label{eq:ConservationOfMomentum}
  N_{1} \Omega_{1}(z) + N_{2} \Omega_{2}(z) = \mathrm{const}.
\end{equation}

@@
